## Algorithmic Trading Strategy

## Introduction

In this assignment, you will develop an algorithmic trading strategy by incorporating financial metrics to evaluate its profitability. This exercise simulates a real-world scenario where you, as part of a financial technology team, need to present an improved version of a trading algorithm that not only executes trades but also calculates and reports on the financial performance of those trades.

## Background

Following a successful presentation to the Board of Directors, you have been tasked by the Trading Strategies Team to modify your trading algorithm. This modification should include tracking the costs and proceeds of trades to facilitate a deeper evaluation of the algorithm’s profitability, including calculating the Return on Investment (ROI).

After meeting with the Trading Strategies Team, you were asked to include costs, proceeds, and return on investments metrics to assess the profitability of your trading algorithm.

## Objectives

1. **Load and Prepare Data:** Open and run the starter code to create a DataFrame with stock closing data.

2. **Implement Trading Algorithm:** Create a simple trading algorithm based on daily price changes.

3. **Customize Trading Period:** Choose your entry and exit dates.

4. **Report Financial Performance:** Analyze and report the total profit or loss (P/L) and the ROI of the trading strategy.

5. **Implement a Trading Strategy:** Implement a trading strategy and analyze the total updated P/L and ROI. 

6. **Discussion:** Summarise your finding.


## Instructions

### Step 1: Data Loading

Start by running the provided code cells in the "Data Loading" section to generate a DataFrame containing AMD stock closing data. This will serve as the basis for your trading decisions. First, create a data frame named `amd_df` with the given closing prices and corresponding dates. 

```{r load-data}

# Load data from CSV file
amd_df <- read.csv("AMD.csv")

# Convert the date column to Date type and Adjusted Close as numeric
amd_df$date <- as.Date(amd_df$Date)
amd_df$close <- as.numeric(amd_df$Adj.Close)

amd_df <- amd_df[, c("date", "close")]
```


##Plotting the Data
Plot the closing prices over time to visualize the price movement.
```{r plot}
plot(amd_df$date, amd_df$close,'l')
```

## Step 2: Customize Trading Period
- Define a trading period you wanted in the past five years 
```{r period}
start_date <- as.Date('2023-01-03')
end_date <- as.Date('2024-01-02')

# Filter the dataframe according to the specified trading period
amd_df <- amd_df[amd_df$date >= start_date & amd_df$date <= end_date, ]

```


## Step 3: Trading Algorithm
Implement the trading algorithm as per the instructions. You should initialize necessary variables, and loop through the dataframe to execute trades based on the set conditions.

- Initialize Columns: Start by ensuring dataframe has columns 'trade_type', 'costs_proceeds' and 'accumulated_shares'.
- Change the algorithm by modifying the loop to include the cost and proceeds metrics for buys of 100 shares. Make sure that the algorithm checks the following conditions and executes the strategy for each one:
  - If the previous price = 0, set 'trade_type' to 'buy', and set the 'costs_proceeds' column to the current share price multiplied by a `share_size` value of 100. Make sure to take the negative value of the expression so that the cost reflects money leaving an account. Finally, make sure to add the bought shares to an `accumulated_shares` variable.
  - Otherwise, if the price of the current day is less than that of the previous day, set the 'trade_type' to 'buy'. Set the 'costs_proceeds' to the current share price multiplied by a `share_size` value of 100.
  - You will not modify the algorithm for instances where the current day’s price is greater than the previous day’s price or when it is equal to the previous day’s price.
  - If this is the last day of trading, set the 'trade_type' to 'sell'. In this case, also set the 'costs_proceeds' column to the total number in the `accumulated_shares` variable multiplied by the price of the last day.



```{r trading}
# NOTE: To customize trading period run {r period} first

# Initialize columns for trade type, cost/proceeds, and accumulated shares in amd_df
amd_df$trade_type <- NA
amd_df$costs_proceeds <- NA  # Corrected column name
amd_df$accumulated_shares <- 0  # Initialize if needed for tracking

# Initialize variables for trading logic
previous_price <- 0
share_size <- 100
accumulated_shares <- 0

for (i in 1:nrow(amd_df)) {
  # Set current price
  current_price <- amd_df$close[i]
  
  # Check if first day
  if (previous_price == 0) {
    amd_df$trade_type[i] <- 'buy'
    amd_df$costs_proceeds[i] <- - share_size * amd_df$close[i]
    accumulated_shares <- share_size
    
  } else if (i == nrow(amd_df)) { # Check if last day
    amd_df$trade_type[i] <- 'sell'
    amd_df$costs_proceeds[i] <- current_price * accumulated_shares
    
  } else if (current_price < previous_price) { # Buy condition
    amd_df$trade_type[i] <- 'buy'
    amd_df$costs_proceeds[i] <- - share_size * amd_df$close[i]
    accumulated_shares <- accumulated_shares + share_size
  }
  
  # Update accumulated shares in the dataframe
  amd_df$accumulated_shares[i] <- accumulated_shares

  # Set previous price for next iteration
  previous_price <- current_price
  
}
```


## Step 4: Run Your Algorithm and Analyze Results
After running your algorithm, check if the trades were executed as expected. Calculate the total profit or loss and ROI from the trades.

- Total Profit/Loss Calculation: Calculate the total profit or loss from your trades. This should be the sum of all entries in the 'costs_proceeds' column of your dataframe. This column records the financial impact of each trade, reflecting money spent on buys as negative values and money gained from sells as positive values.
- Invested Capital: Calculate the total capital invested. This is equal to the sum of the 'costs_proceeds' values for all 'buy' transactions. Since these entries are negative (representing money spent), you should take the negative sum of these values to reflect the total amount invested.
- ROI Formula: $$\text{ROI} = \left( \frac{\text{Total Profit or Loss}}{\text{Total Capital Invested}} \right) \times 100$$

```{r}
# For the basic algorithm:

# Sum the cost proceeds column whilst ignoring all the NA entries
total_profit <- sum(amd_df$costs_proceeds, na.rm = TRUE)
cat("Total profit is", total_profit, "\n")

# Sum the cost proceeds column if the value is negative whilst ignoring all the NA entries
total_capital_invested <- -sum(amd_df$costs_proceeds[amd_df$costs_proceeds < 0], na.rm = TRUE)
cat("Total captial invested is", total_capital_invested, "\n")

# Calcuate ROI using the formula
roi <- total_profit/total_capital_invested * 100
cat("ROI is", roi, "%\n")

```

## Step 5: Profit-Taking Strategy or Stop-Loss Mechanisum (Choose 1)
- Option 1: Implement a profit-taking strategy that you sell half of your holdings if the price has increased by a certain percentage (e.g., 20%) from the average purchase price.
- Option 2: Implement a stop-loss mechanism in the trading strategy that you sell half of your holdings if the stock falls by a certain percentage (e.g., 20%) from the average purchase price. You don't need to buy 100 stocks on the days that the stop-loss mechanism is triggered.

```{r option}
# Option 1: Profit-Taking Strategy

# Initialize columns for trade type, cost/proceeds, and accumulated shares in amd_df
amd_df$trade_type <- NA
amd_df$costs_proceeds <- NA  # Corrected column name
amd_df$accumulated_shares <- 0  # Initialize if needed for tracking

# Initialize variables for trading logic
previous_price <- 0
share_size <- 100
accumulated_shares <- 0

avg_price <- 0
amd_df$avg_price <- 0
take_profit_percentage <- 1.2

for (i in 1:nrow(amd_df)) {
  # Set current price
  current_price <- amd_df$close[i]
  
  # Check if first day
  if (previous_price == 0) {
    amd_df$trade_type[i] <- 'buy'
    amd_df$costs_proceeds[i] <- - share_size * amd_df$close[i]
    accumulated_shares <- share_size
    avg_price <- current_price
    amd_df$avg_price[i] <- avg_price
    
  } else if (i == nrow(amd_df)) { # Check if last day
    amd_df$trade_type[i] <- 'sell'
    amd_df$costs_proceeds[i] <- current_price * accumulated_shares
    
  } else if (current_price > take_profit_percentage * avg_price) { # Sell condition
    # Sell half your shares
    amd_df$trade_type[i] <- 'sell'
    amd_df$costs_proceeds[i] <- accumulated_shares / 2 * current_price
    accumulated_shares <- accumulated_shares / 2
    
  } else if (current_price < previous_price) { # Buy condition
    amd_df$trade_type[i] <- 'buy'
    amd_df$costs_proceeds[i] <- - share_size * amd_df$close[i]
    accumulated_shares <- accumulated_shares + share_size
    
    # Calculate average purchase price
    avg_price <- (avg_price * (accumulated_shares - share_size) + (current_price*share_size)) / accumulated_shares
  }
  
  # Update accumulated shares in the dataframe
  amd_df$accumulated_shares[i] <- accumulated_shares

  # Set previous price for next iteration
  previous_price <- current_price
  
  # Update average purchase price in the dataframe
  amd_df$avg_price[i] <- avg_price
  
# Fill your code here
}
```


## Step 6: Summarize Your Findings
- Did your P/L and ROI improve over your chosen period?
- Relate your results to a relevant market event and explain why these outcomes may have occurred.


```{r}
# For the Profit Taking algorithm:

# Sum the cost proceeds column whilst ignoring all the NA entries
total_profit <- sum(amd_df$costs_proceeds, na.rm = TRUE)
cat("Total profit is", total_profit, "\n")

# Sum the cost proceeds column if the value is negative whilst ignoring all the NA entries
total_capital_invested <- -sum(amd_df$costs_proceeds[amd_df$costs_proceeds < 0], na.rm = TRUE)
cat("Total captial invested is", total_capital_invested, "\n")

# Calcuate ROI using the formula
roi <- total_profit/total_capital_invested * 100
cat("ROI is", roi, "%\n")

```

The P/L and ROI for the original algorithm was $428886 and 35.96529 % whilst my P/L and ROI for the profit-taking strategy was $256399 and 26.70439 % meaning my ROI for the profit-taking strategy was lower and did not improve compared to the original algorithm. This is because the profit-taking strategy provides a risk-managed approach, aiming to secure gains by selling positions incrementally as they reach predefined profit targets. While this approach can lead to a more stable portfolio and protect against market downturns, it may result in missing out on potential gains in a strongly bullish market compared to a buy and hold strategy.


During my selected trading period (2023-01-03 to 2024-01-02), a key market event was the 14th of March 2023 where the announcement of the 4th Gen AMD EPYC™ Processors which featured world class performance and energy efficiency caused an increase in investor confidence. The rise in AMD shares on the following Thursday (March 16) signals high investor confidence in the company's potential, with share price jumping from to $87.45 on the 14th to $96.60 on the 16th. This significant price increase reflects the market's positive reaction to AMD's technological advancements and its strong positioning in the semiconductor industry. This strong performance of AMD is reflected in our ROI being higher in our buy and hold strategy as we are able to capture the full upside potential of the company whilst our ROI for the profit taking strategy is lower due to us selling our shares prematurely and not fully benefiting from the price surge following positive market events.


Additionally, I found that we can improve our ROI using the profit taking method by setting a higher profit taking percentage as we get an ROI of 31.1383 % using a take profit percentage of 40 % as opposed to 20 %. This is because we lock in profits when price reaches a higher threshold, allowing us to capture a higher return in a strongly performing market, however having a higher threshold may put us at risk of a market downturn before reaching our profit target. To further reduce our risk in a volatile market we can also consider adding a stop loss algorithm simultaneously which protects us from large losses if the market performs poorly.

In conclusion, our ROI for the profit taking strategy was lower than our original buy and hold strategy as AMD performed strongly and we were not able to capture the full upside using the profit taking strategy.
